Model Checking Differentially Private Properties

### 2018-06-07 above-threshold.tex
Write something on experiments
polish and spell check

Below we describe an online mechanism from~\cite{DR:14:AFDP}.
Given a threshold and a series of adaptive queries, we care for
the queries whose results are above the threshold; queries below the
threshold only disclose minimal information and hence are considered
irrelevant. Since query results above the threshold convey useful
information, let us assume the mechanism will halt on the first such
query result for simplicity. 
%The question is to design a privacy-preserving mechanism to have 
%similar behaviors on similar databases. 
In~\cite{DR:14:AFDP}, a mechanism is design for continuous
queries by applying the Laplace mechanism. We will develop a mechanism
for bounded and discrete queries using the geometric mechanism.

@@

	\end{algorithmic}
	\caption{Input: private database $d$, 
          queries $f_i : \{ 0, 1, \ldots, 5 \} \rightarrow \{ 0, 1,
          \ldots, 5 \}$ with sensitivity $1$, threshold $t \in \{ 0,
          1, \ldots, 5 \}$; Output: $a_1, a_2, \ldots$}
	\label{algorithm:online-model}
\end{algorithm}

Consider a threshold $t \in
\{ 0, 1 \ldots, 5\}$, and queries $\{ f_i : f_i $ with sensitivity $\Delta
(f_i) = 1 \}$. In order to protect privacy, our mechanism applies the
truncated $\frac{1}{4}$-geometric mechanism to obtain a perturbed
threshold $t'$. For each query, the truncated $\frac{1}{2}$-geometric
mechanism is applied to its result. If the perturbed result is not less
than the perturbed threshold, our mechanism outputs $\top$ and
halts. Otherwise, it outputs $\bot$ and continues to the next query
(Algorithm~\ref{algorithm:online-model}). The above threshold
mechanism outputs a sequence of the form $\bot^* \top$. On
similar databases, we would like to show that the above
threshold mechanism outputs the same sequence with similar
probabilities.

@@

It is not hard to model the above threshold mechanism as a Markov
decision process (Figure~\ref{figure:mdp-above-threshold}). In the
figure, we sketch the model where the threshold and query results are
in $\{ 0, 1, 2 \}$. The model simulates two computation in parallel:
one for the dataset, the other for its neighbor. The state $t_ir_j$
represents the input threshold $i$ and the first query result
$j$; the state $t'_ir'_j$ represents the perturbed threshold $i$ and
the perturbed query result $j$. Other states are similar. Consider the
state $t_0r_1$. After applying the truncated $\frac{1}{4}$-geometric
mechanism, it goes to one of the states $t'_0r_1$, $t'_1r_1$
accordingly. From the state $t'_1r_1$, for instance, it moves to one
of $t'_1r'_0$, $t'_1r'_1$, $t'_1r'_2$ by applying the truncated
$\frac{1}{2}$-geometric mechanism to the query result. If it arrives
at $t'_1r'_1$ or $t'_1r'_2$, the perturbed query result is not less
than the perturbed threshold. The model enters the state of self
loop. Otherwise, it moves to one of $t'_1r_0$, $t'_1r_1$, or $t'_1r_2$
by actions non-deterministically (double arrows). The computation of
the neighbor is similar. We just use the underlined symbols to
represent threshold and query results. For instance, the state
$\underline{t}'_2\underline{r}'_1$ represents the perturbed threshold
$2$ and the perturbed query result $1$ in the neighbor.

Now, the non-deterministic choices in the two computation cannot be 
independent. Recall that the sensitivity of each query $f_i$ is
$1$. If the top computation moves to the state, say, $t'_1r_0$, it
means the next query result on the dataset is $0$. Subsequently, the
bottom computation can only move to $\underline{t}'_i\underline{r}_0$
or $\underline{t}'_i\underline{r}_1$ depending on the perturbed
threshold. This is where action symbols are useful. Define the action
symbols $\{ m\underline{n} : |m - n| \leq 1 \}$. The action
$m\underline{n}$ represents that the next query results for the dataset
and its neighbor are $m$ and $n$ respectively. For instance, the
non-deterministic choice from $t'_1r'_0$ to $t'_1r_0$ is associated
with two actions $0\underline{0}$ and $0\underline{1}$ (but not
$0\underline{2}$). Similarly, the choice from
$\underline{t}'_2\underline{r}'_1$ to
$\underline{t}'_2\underline{r}_0$ is associated with the actions
$0\underline{0}$ and $1\underline{0}$ (but not $2\underline{0}$). On
action $0\underline{0}$, both computation will move to $t'_ir_0$ and
$\underline{t}'_j\underline{r}_0$ respectively by the perturbed
thresholds. Action symbols make sure the two computation on neighbor
datasets are modeled properly. It remains to define the neighborhood
relation. Since the sensitivity is $1$, it suffices to consider 
$\{ (t_ir_m, t_ir_m),
    (\underline{t}_i\underline{r}_n, \underline{t}_i\underline{r}_n),
    (t_ir_m, \underline{t}_i\underline{r}_n), 
    (\underline{t}_i\underline{r}_n, t_ir_m) : | m - n | \leq 1 \}$.

>>>>

Below we describe an online mechanism from~\cite{DR:14:AFDP}.
Given a threshold and a series of adaptive queries, we care for
the queries whose results are above the threshold; queries below the
threshold only disclose minimal information and hence 
irrelevant. Since query results above the threshold convey useful
information, let us assume the mechanism will halt on the first such
query result for simplicity. 
%The question is to design a privacy-preserving mechanism to have 
%similar behaviors on similar databases. 
In~\cite{DR:14:AFDP}, a mechanism is designed for continuous
queries by applying the Laplace mechanism. We will develop a mechanism
for discrete bounded queries using the truncated geometric mechanism.

@@

	\end{algorithmic}
	\caption{Input: private database $d$, 
          queries $f_i : d \rightarrow \{ 0, 1,
          \ldots, 5 \}$ with sensitivity $1$, threshold $t \in \{ 0,
          1, \ldots, 5 \}$; Output: $a_1, a_2, \ldots$}
	\label{algorithm:online-model}
\end{algorithm}

Assume that we have a threshold $t \in \{ 0, 1, \ldots, 5 \}$
and queries $\{ f_i : f_i $ with sensitivity $\Delta
(f_i) = 1 \}$. In order to protect privacy, our mechanism applies the
truncated $\frac{1}{4}$-geometric mechanism to obtain a perturbed
threshold $t'$. For each query $f_i$, the truncated $\frac{1}{2}$-geometric
mechanism is applied to its result $r_i$. If the perturbed result
$r'_i$ is not less
than the perturbed threshold $t'$, the mechanism halts with the input
$\top$. Otherwise, it outputs $\bot$ and continues to the next query
(Algorithm~\ref{algorithm:online-model}). The above threshold
mechanism outputs a sequence of the form $\bot^* \top$. On
similar datasets, we want to show that the above
threshold mechanism outputs the same sequence with similar
probabilities.

@@

It is not hard to model the above threshold mechanism as a Markov
decision process (Figure~\ref{figure:mdp-above-threshold}). In the
figure, we sketch the model where the threshold and query results are
in $\{ 0, 1, 2 \}$. The model simulates two computation in parallel:
one for the dataset, the other for its neighbor. The state $t_ir_j$
represents the input threshold $i$ and the first query result
$j$; the state $t'_ir'_j$ represents the perturbed threshold $i$ and
the perturbed query result $j$. Other states are similar. Consider the
state $t_0r_1$. After applying the truncated $\frac{1}{4}$-geometric
mechanism, it goes to one of the states $t'_0r_1$, $t'_1r_1$, $t'_2r_1$
accordingly. From the state $t'_1r_1$, for instance, it moves to one
of $t'_1r'_0$, $t'_1r'_1$, $t'_1r'_2$ by applying the truncated
$\frac{1}{2}$-geometric mechanism to the query result. If it arrives
at $t'_1r'_1$ or $t'_1r'_2$, the perturbed query result is not less
than the perturbed threshold. The model halts with the output $\top$
by entering the state of a self loop. 
Otherwise, it moves to one of $t'_1r_0$, $t'_1r_1$, or $t'_1r_2$
non-deterministically (double arrows). The computation of
its neighbor is similar. We just use the underlined symbols to
represent threshold and query results. For instance, the state
$\underline{t}'_2\underline{r}'_1$ represents the perturbed threshold
$2$ and the perturbed query result $1$ in the neighbor.

Now, the non-deterministic choices in the two computation cannot be 
independent. Recall that the sensitivity of each query $f_i$ is
$1$. If the top computation moves to the state, say, $t'_1r_0$, it
means the next query result on the dataset is $0$. Subsequently, the
bottom computation can only move to $\underline{t}'_i\underline{r}_0$
or $\underline{t}'_i\underline{r}_1$ depending on the perturbed
threshold. This is where actions are useful. Define the actions
$\{ m\underline{n} : |m - n| \leq 1 \}$. The action
$m\underline{n}$ represents that the next query result for the dataset
and its neighbor are $m$ and $n$ respectively. For instance, the
non-deterministic choice from $t'_1r'_0$ to $t'_1r_0$ is associated
with two actions $0\underline{0}$ and $0\underline{1}$ (but not
$0\underline{2}$). Similarly, the choice from
$\underline{t}'_2\underline{r}'_1$ to
$\underline{t}'_2\underline{r}_0$ is associated with the actions
$0\underline{0}$ and $1\underline{0}$ (but not $2\underline{0}$). On
the action $0\underline{0}$, the top computation moves to $t'_ir_0$ and
the bottom computation moves to $\underline{t}'_j\underline{r}_0$ by
their respective perturbed thresholds.
Actions make sure the two computation on neighbors
is modeled properly. It remains to define the neighborhood
relation. Recall the sensitivity is $1$. Consider the neighborhood
relation 
$\{ (t_ir_m, t_ir_m),
    (\underline{t}_i\underline{r}_n, \underline{t}_i\underline{r}_n),
    (t_ir_m, \underline{t}_i\underline{r}_n), 
    (\underline{t}_i\underline{r}_n, t_ir_m) : | m - n | \leq 1 \}$.
That is, two states are neighbors if they represent two inputs of the
same threshold and query results with difference at most one.

### 2018-06-07 examples.tex
reorganize sections
polish and spell check

To model differentially private mechanisms by Markov chains, we
formalize inputs (such as databases or query results) as states. 
Randomized computation is modeled by probabilistic transitions. Atomic
propositions are used to designate intended interpretation on states
(such as inputs or outputs). We demonstrates the ideas in examples.

@@

Consider the survey question: have you been diagnosed
with the disease $X$? In order to protect surveyees' privacy, each
answers the question as follows. The surveyee first flips a
coin. If it is tail, she answers the question truthfully. Otherwise,
she randomly answers \textit{Yes} or \textit{No}
uniformly (Figure~\ref{figure:2-dp-table})~\cite{DR:14:AFDP}.

Let us analyze the mechanism briefly. The data universe $\mathcal{X}$
is $\{ +, - \}$. The mechanism $M$ is a randomized algorithm with
inputs in $\mathcal{X}$ and outputs in $\{ Y, N \}$. For any $x \in
\mathcal{X}$, we have $\frac{1}{4} \leq \Pr[M (x) = Y] \leq
\frac{3}{4}$. Hence $\Pr[M (x) = Y] \leq \frac{3}{4} = 3 \cdot
\frac{1}{4} \leq e^{\ln 3} \Pr[M (x') = Y]$ for any neighbors $x, x'
\in \mathcal{X}$. Similarly, $\Pr[M (x) = N] \leq e^{\ln 3} \Pr[M (x')
= N]$. The survey mechanism is $(\ln 3, 0)$-differentially private.
The random noise boosts the probability of answering
\textit{Yes} or \textit{No} to at least $\frac{1}{4}$ regardless of
diagnoses. Inferences on individual diagnosis can be plausibly denied.

\begin{figure}
  \centering
  \begin{subfigure}{.48\columnwidth}
      \[
      \begin{array}{|c|c|c|}
        \hline
        &
        \multicolumn{2}{c|}{output}
        \\
        \hline
        state & Y & N \\
        \hline
        + & \frac{3}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 1
          & \frac{1}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 0
        \\
        \hline
        - & \frac{1}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 0
          & \frac{3}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 1
        \\
        \hline
      \end{array}
      \]
    \caption{Survey Mechanism}
    \label{figure:2-dp-table}
  \end{subfigure}
  \hspace{.05\columnwidth}
  \begin{subfigure}{.40\columnwidth}
    \resizebox{\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (p) at ( 0,  .5) { $+$ };
      \node[node] (n) at ( 0, -.5) { $-$ };
      \node[node] (Y) at (-1.5,  0) { $s$ };
      \node at (-1.5, .5) { $\mathit{out}_Y$ };
      \node[node] (N) at ( 1.5,  0) { $t$ };
      \node at ( 1.5, .5) { $\mathit{out}_N$ };

@@

\subsection{Truncated $\alpha$-Geometric Mechanism}\label{subsec:geometric}
More sophisticated differentially private mechanisms are
available. Consider a query
$f : \mathcal{X}^n \rightarrow \{ 0, 1, \ldots, m \}$. Let $\alpha \in (0, 1)$.
The \emph{$\alpha$-geometric mechanism}
outputs $f(x) + Y$ where $Y$ is a random variable with the geometric
distribution~\cite{GRS:09:UUPM,GRS:12:UUPM} :
\[
\Pr[Y = y] = \frac{1 - \alpha}{1 + \alpha}\alpha^{|y|}
\textmd{ for } y \in \bbfZ.
\]
%Observe that $\Pr[Y=0] \rightarrow 1$ as $\alpha \rightarrow 0$. The
%$\alpha$-geometric mechanism is less noisy for smaller $\alpha$.
The $\alpha$-geometric mechanism is oblivious since it has the same
output distribution on any inputs $x, x'$ with $f (x) = f
(x')$. It is $(- {\Delta (f)} \ln \alpha, 0)$-differentially
private for any query $f$ with sensitivity $\Delta (f)$. Observe that
the privay guarantee $(-{\Delta (f)} \ln \alpha, 0)$ depends on the
sensitivity of the query $f$. To achieve $(\epsilon, 0)$-differential
privacy using the $\alpha$-geometric mechanism, one first decides the
sensitivity of the query and then computes the parameter $\alpha =
e^{-{\epsilon}/{{\Delta (f)}}}$.

@@

\begin{figure}[tbh]
  \centering
  \begin{subfigure}{.40\columnwidth}
    \[
    \begin{array}{|c|c|c|c|c|c|c|}
      \hline
      \textmd{input/output} & 0 & 1 & 2 & 3 & 4 & 5\\
      \hline
      0 & {2}/{3}  & {1}/{6} & {1}/{12} & {1}/{24}  & {1}/{48} & {1}/{48}\\
      \hline
      1 & {1}/{3}  & {1}/{3} & {1}/{6} & {1}/{12}  & {1}/{24} & {1}/{24}\\
      \hline
      2 & {1}/{6}  & {1}/{6} & {1}/{3} & {1}/{6}   & {1}/{12} & {1}/{12}\\
      \hline
      3 & {1}/{12} & {1}/{12} & {1}/{6} & {1}/{3}   & {1}/{6} & {1}/{6}\\
      \hline
      4 & {1}/{24} & {1}/{24} & {1}/{12} & {1}/{6}   & {1}/{3} & {1}/{3}\\
      \hline
      5 & {1}/{48} & {1}/{48} & {1}/{24} & {1}/{12} & {1}/{6} & {2}/{3}\\
      \hline
    \end{array}
    \]
    \caption{$\frac{1}{2}$-Geometric Mechanism}
    \label{figure:geometric-mechanism-table}
  \end{subfigure}
  \hspace{.08\columnwidth}
  \begin{subfigure}{.50\columnwidth}
    \centering
  \resizebox{.85\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (i0) at (0,  1) { $s_0$ };
      \node at (0, 1.5) { $\mathit{in}_0$ };
      \node at (0, 0) { $\vdots$ };
      \node at (.8, -.8) { $\vdots$ };
      \node[node] (i5) at (0, -1) { $s_5$ };
      \node at (0, -.5) { $\mathit{in}_5$ };

@@

Similar to the survey mechanism, it is straightforward to model the
truncated $\frac{1}{2}$-geometric mechanism as a Markov chain. 
Recall that the truncated $\frac{1}{2}$-geometric mechanism is
oblivious. It suffices to consider $\{ f (x) : x \in \mathcal{X}^n \}$
instead of $\mathcal{X}^n$ as inputs. 
Let the state $s_k$ and $t_l$
denote the input value $k$ and output value $l$ respectively. Define
the state set $S = \{ s_k, t_k : k \in \{ 0, 1, \ldots, m \} \}$.
The probability transition $\wp (s_k, t_l)$ is the probability of the
output value $l$ on the input value $k$ as defined in the
mechanism. Moreover, we have $\mathit{in}_k \in L (s_k)$ and
$\mathit{out}_k \in L (t_k)$ for $k \in \{ 0, 1, \ldots, n \}$.
If $\Delta (f) = 1$, $| f (x) - f (x') | \leq 1$ for every neighbors
$x, x' \in \mathcal{X}^n$. Subsequently, $s_k$ and $s_l$ are
neighbors iff $| k - l | \leq 1$ in our model.
Figure~\ref{figure:geometric-mechanism-markov-chain} gives
the Markov chain for the truncated
$\frac{1}{2}$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$.

\subsection{Subsampling Majority}
\label{subsection:subsampling}

\input{subsampling}

>>>>

To model differentially private mechanisms by Markov chains, we
formalize inputs (such as datasets or query results) as states. 
Randomized computation is modeled by probabilistic transitions. Atomic
propositions are used to designate intended interpretation on states
(such as inputs or outputs). We demonstrates these ideas in examples.

@@

Consider the survey question: have you been diagnosed
with the disease $X$? In order to protect surveyees' privacy, each
answers the question as follows. The surveyee first flips a
coin. If it is tail, she answers the question truthfully. Otherwise,
she randomly answers $1$ or $0$
uniformly (Figure~\ref{figure:2-dp-table})~\cite{DR:14:AFDP}.

Let us analyze the mechanism briefly. The data universe $\mathcal{X}$
is $\{ +, - \}$. The mechanism $M$ is a randomized algorithm with
inputs in $\mathcal{X}$ and outputs in $\{ 1, 0 \}$. For any $x \in
\mathcal{X}$, we have $\frac{1}{4} \leq \Pr[M (x) = 1] \leq
\frac{3}{4}$. Hence $\Pr[M (x) = 1] \leq \frac{3}{4} = 3 \cdot
\frac{1}{4} \leq e^{\ln 3} \Pr[M (x') = 1]$ for any neighbors $x, x'
\in \mathcal{X}$. Similarly, $\Pr[M (x) = 0] \leq e^{\ln 3} \Pr[M (x')
= 0]$. The survey mechanism is $(\ln 3, 0)$-differentially private.
The random noise boosts the probability of answering
$1$ or $0$ to at least $\frac{1}{4}$ regardless of
diagnoses. Inferences on individual diagnosis can be plausibly denied.

\begin{figure}
  \centering
  \begin{subfigure}{.48\columnwidth}
      \[
      \begin{array}{c|c|c|c|}
        \multicolumn{2}{c}{} &
        \multicolumn{2}{c}{output}
        \\
        \cline{2-4}
        &  & 1 & 0 \\
        \cline{2-4}
        \multirow{2}*{\rotatebox{90}{state}} 
        &
        + & \frac{3}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 1
          & \frac{1}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 0
        \\
        \cline{2-4}
        &
        - & \frac{1}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 0
          & \frac{3}{4} = \frac{1}{2} \cdot \frac{1}{2} + \frac{1}{2} \cdot 1
        \\
        \cline{2-4}
      \end{array}
      \]
    \caption{Survey Mechanism}
    \label{figure:2-dp-table}
  \end{subfigure}
  \hspace{.05\columnwidth}
  \begin{subfigure}{.40\columnwidth}
    \resizebox{\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (p) at ( 0,  .5) { $+$ };
      \node[node] (n) at ( 0, -.5) { $-$ };
      \node[node] (Y) at (-1.5,  0) { $s$ };
      \node at (-1.5, .5) { $\mathit{out}_1$ };
      \node[node] (N) at ( 1.5,  0) { $t$ };
      \node at ( 1.5, .5) { $\mathit{out}_0$ };

@@

\subsection{Truncated $\alpha$-Geometric Mechanism}\label{subsec:geometric}
More sophisticated differentially private mechanisms are
available. Consider a query
$f : \mathcal{X}^n \rightarrow \{ 0, 1, \ldots, m \}$. Let $\alpha \in (0, 1)$.
The \emph{$\alpha$-geometric mechanism}
outputs $f(x) + Y$ where $Y$ is a random variable with the geometric
distribution~\cite{GRS:09:UUPM,GRS:12:UUPM} :
\[
\Pr[Y = y] = \frac{1 - \alpha}{1 + \alpha}\alpha^{|y|}
\textmd{ for } y \in \bbfZ.
\]
%Observe that $\Pr[Y=0] \rightarrow 1$ as $\alpha \rightarrow 0$. The
%$\alpha$-geometric mechanism is less noisy for smaller $\alpha$.
The $\alpha$-geometric mechanism is oblivious since it has the same
output distribution on any inputs $x, x'$ with $f (x) = f
(x')$. It is $(- {\Delta (f)} \ln \alpha, 0)$-differentially
private for any query $f$ with sensitivity $\Delta (f)$. Observe that
the privacy guarantee $(-{\Delta (f)} \ln \alpha, 0)$ depends on the
sensitivity of the query $f$. To achieve $(\epsilon, 0)$-differential
privacy using the $\alpha$-geometric mechanism, one first decides the
sensitivity of the query and then computes the parameter $\alpha =
e^{-{\epsilon}/{{\Delta (f)}}}$.

@@

\begin{figure}[tbh]
  \centering
  \begin{subfigure}{.40\columnwidth}
    \[
    \begin{array}{c|c|c|c|c|c|c|c|}
      \multicolumn{2}{c}{} &
      \multicolumn{6}{c}{output} \\
      \cline{2-8}
      & & 0 & 1 & 2 & 3 & 4 & 5\\
      \cline{2-8}
      \multirow{6}*{\rotatebox{90}{input}}
      & 0 & {2}/{3}  & {1}/{6} & {1}/{12} & {1}/{24}  & {1}/{48} & {1}/{48}\\
      \cline{2-8}
      & 1 & {1}/{3}  & {1}/{3} & {1}/{6} & {1}/{12}  & {1}/{24} & {1}/{24}\\
      \cline{2-8}
      & 2 & {1}/{6}  & {1}/{6} & {1}/{3} & {1}/{6}   & {1}/{12} & {1}/{12}\\
      \cline{2-8}
      & 3 & {1}/{12} & {1}/{12} & {1}/{6} & {1}/{3}   & {1}/{6} & {1}/{6}\\
      \cline{2-8}
      & 4 & {1}/{24} & {1}/{24} & {1}/{12} & {1}/{6}   & {1}/{3} & {1}/{3}\\
      \cline{2-8}
      & 5 & {1}/{48} & {1}/{48} & {1}/{24} & {1}/{12} & {1}/{6} & {2}/{3}\\
      \cline{2-8}
    \end{array}
    \]
    \caption{$\frac{1}{2}$-Geometric Mechanism}
    \label{figure:geometric-mechanism-table}
  \end{subfigure}
  \hspace{.08\columnwidth}
  \begin{subfigure}{.50\columnwidth}
    \centering
  \resizebox{.85\columnwidth}{!}{
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node
      distance=2cm,node/.style={circle,draw}]
      \node[node] (i0) at (0,  1) { $s_0$ };
      \node at (0, 1.5) { $\mathit{in}_0$ };
      \node at (0, 0) { $\vdots$ };
      \node at (.8, -.8) { $\vdots$ };
      \node[node] (i5) at (0, -1) { $s_5$ };
      \node at (0, -.5) { $\mathit{in}_5$ };

@@

Similar to the survey mechanism, it is straightforward to model the
truncated $\frac{1}{2}$-geometric mechanism as a Markov chain. 
Recall that the truncated $\frac{1}{2}$-geometric mechanism is
oblivious. It suffices to consider $\{ f (x) : x \in \mathcal{X}^n \}$
as inputs. 
Let the state $s_k$ and $t_l$
denote the input $k$ and output $l$ respectively. Define
$S = \{ s_k, t_k : k \in \{ 0, 1, \ldots, m \} \}$.
The probability transition $\wp (s_k, t_l)$ is the probability of the
output $l$ on the input $k$ as defined in the
mechanism. Moreover, we have $\mathit{in}_k \in L (s_k)$ and
$\mathit{out}_k \in L (t_k)$ for $k \in \{ 0, 1, \ldots, n \}$.
If $\Delta (f) = 1$, $| f (x) - f (x') | \leq 1$ for every neighbors
$x, x' \in \mathcal{X}^n$. Subsequently, $s_k$ and $s_l$ are
neighbors iff $| k - l | \leq 1$ in our model.
Figure~\ref{figure:geometric-mechanism-markov-chain} gives
the Markov chain for the truncated
$\frac{1}{2}$-geometric mechanism over $\{ 0, 1, \ldots, 5 \}$.

### 2018-06-07 introduction.tex
reorganize sections
polish and spell check

Differential privacy formalizes the tradeoff between privacy and
utility in data analysis. Intuitively, a randomized data analysis
mechanism is differentially private if it behaves similarly on similar
input databases~\cite{DMNS:06:CNSPD,D:06:DP}. Consider, for example,
the Laplace mechanism where analysis results are perturbed by random
noises with the Laplace distribution~\cite{DR:14:AFDP}. Random noises
hide the differences of analysis results from similar databases.
Clearly, more
randomness gives more privacy but less utility in released perturbed
results. Under the framework of differential privacy, data analysts
can balance the tradeoff rigorously in their data analysis
mechanisms~\cite{DR:14:AFDP,JLE:14:DPML}.

@@

Designing differentially private mechanisms can be tedious for
sophisticated data analyses. Privacy leak has also been observed in
data analysis programs implementing differentialy
privacy~\cite{M:12:SLSBDP,TKBW:17:PLAI}. This calls for formal 
analysis of differential privacy on both designs and implementations.
In this paper, we propose the logic $\dpCTLstar$ for specifying
differential privacy and investigate their model checking
problems. Data analysts can automatically verify their designs and
implementations with our techniques. Most interestingly, our
techniques can be adopted easily by existing probabilistic
model checkers. Privacy checking with existing tools is attainable
with minimal efforts. More interaction
between model checking~\cite{ManneP92,BK:08:PMC} and privacy analysis hopefully will follow.

@@

In order to illustrate applicability of our techniques, we give
detailed formalizations of several data analysis mechanisms in this
paper.
In differential privacy, data analysis mechanisms are
but randomized algorithms. We follow the standard practice in
probabilistic model checking to formalize such mechanisms
by Markov chains or Markov decision processes~\cite{Put05}. We give
formalizations of several differentially private mechanisms and
explain subtle privacy conditions.
When a data analysis mechanism does not
interact with its environment, it is formalized as a Markov
chain. Otherwise, its interactions are formalized by actions in
Markov decision processes. Our formalization effectively assumes that
actions are controlled by adversaries. It is thus necessary to
consider all privacy attacks from adversaries in order to establish
differential privacy.
%\todo{What is a state?}

@@

Two ingredients are introduced to specify differentially private
behaviors. A reflexive and commutative user-defined binary relation
over states is required to formalize similarity. We moreover add
the path quantifier $\dpriv{\epsilon}{\delta}$ for similar
behaviors. Informally, a state satisfies $\dpriv{\epsilon}{\delta}
\phi$ if it has probabilistically similar path behavior $\phi$ as
its similar states. Consider, for instance, a binary data analysis
mechanism computing the likelihood ($\mathit{high}$ or $\mathit{low}$)
of an epidemic in a population. Then a state satisfies
$\dpriv{\epsilon}{\delta} (\F \mathit{high}) \wedge
\dpriv{\epsilon}{\delta} (\F \mathit{low})$ if the mechanism is
$(\epsilon, \delta)$-differentially private at the state.
%\todo{not clear here}\lz{here it should be $\X$?}

% differential privacy of infinite behaviors

An important byproduct of this work is to analyze privacy over
infinite behaviors. Consider the formula $\dpriv{\epsilon}{\delta}
(\G \F \mathit{high})$. A state satisfies the formula if it
is as likely to answer $\mathit{high}$ infinitely often
as similar states. Compared to the unbounded
but finite behavior specified by $\F \mathit{high}$,
$\G \F \mathit{high}$ specifies an infinite behavior.
Not only can differential privacy specify randomized algorithms, it
also generalizes to reactive systems naturally.
More data analysis mechanisms are admitable for sophisticated formal
privacy analysis through our techniques.

@@

\noindent
\paragraph{Contributions.} Our main contributions are threefold. 
\begin{enumerate}
\item We introduce the logic $\dpCTLstar$ for reasoning about properties in
differential privacy. The logic is able to to express classical
differentially private properties. Complex properties such as those
involving infinite behaviors can also be expressed;
\item We model several differentially private mechanisms, including stream processing, in Markov chains or Markov decision processes; and
\item We show that the model checking
problem for Markov chains is standard. For Markov decision processes, we show that it is undecidable.
\end{enumerate}

>>>>

Differential privacy formalizes the tradeoff between privacy and
utility in data analysis. Intuitively, a randomized data analysis
mechanism is differentially private if it behaves similarly on similar
input datasets~\cite{DMNS:06:CNSPD,D:06:DP}. Consider, for example,
the Laplace mechanism where analysis results are perturbed by random
noises with the Laplace distribution~\cite{DR:14:AFDP}. Random noises
hide the differences of analysis results from similar datasets.
Clearly, more noises
give more privacy but less utility in released perturbed
results. Under the framework of differential privacy, data analysts
can balance the tradeoff rigorously in their data analysis
mechanisms~\cite{DR:14:AFDP,JLE:14:DPML}.

@@

Designing differentially private mechanisms can be tedious for
sophisticated data analyses. Privacy leak has also been observed in
data analysis programs implementing differential
privacy~\cite{M:12:SLSBDP,TKBW:17:PLAI}. This calls for formal 
analysis of differential privacy on both designs and implementations.
In this paper, we propose the logic $\dpCTLstar$ for specifying
differential privacy and investigate their model checking
problems. Data analysts can automatically verify their designs and
implementations with our techniques. Most interestingly, our
techniques can be adopted easily by existing probabilistic
model checkers. Privacy checking with existing tools is attainable
with minimal efforts. More interaction
between model checking~\cite{ManneP92,BK:08:PMC} and privacy analysis hopefully will follow.

@@

In order to illustrate applicability of our techniques, we give
detailed formalizations of several data analysis mechanisms in this
paper.
In differential privacy, data analysis mechanisms are
but randomized algorithms. We follow the standard practice in
probabilistic model checking to formalize such mechanisms
as Markov chains or Markov decision processes~\cite{Put05}. 
%We give formalizations of several differentially private mechanisms
%and explain subtle privacy conditions.
When a data analysis mechanism does not
interact with its environment, it is formalized as a Markov
chain. Otherwise, its interactions are formalized by actions in
Markov decision processes. Our formalization effectively assumes that
actions are controlled by adversaries. It is thus necessary to
consider all privacy attacks from adversaries in order to establish
differential privacy.
%\todo{What is a state?}

@@

Two ingredients are introduced to specify differentially private
behaviors. A reflexive and symmetric user-defined binary relation
over states is required to formalize similarity. We moreover add
the path quantifier $\dpriv{\epsilon}{\delta}$ for specifying similar
behaviors. Informally, a state satisfies $\dpriv{\epsilon}{\delta}
\phi$ if its probability of having path $\phi$ is close to
those of similar states. Consider, for instance, a binary data analysis
mechanism computing the likelihood ($\mathit{high}$ or $\mathit{low}$)
of an epidemic. Then a state satisfying
$\dpriv{\epsilon}{\delta} (\F \mathit{high}) \wedge
\dpriv{\epsilon}{\delta} (\F \mathit{low})$ denotes similar states
have the same outcomes with similar probabilities.
%\todo{not clear here}\lz{here it should be $\X$?}

\hide{
% differential privacy of infinite behaviors

An important byproduct of this work is to analyze privacy over
infinite behaviors. Consider the formula $\dpriv{\epsilon}{\delta}
(\G \F \mathit{high})$. A state satisfies the formula if it
is as likely to answer $\mathit{high}$ infinitely often
as similar states. Compared to the unbounded
but finite behavior specified by $\F \mathit{high}$,
$\G \F \mathit{high}$ specifies an infinite behavior.
Not only can differential privacy specify randomized algorithms, it
also generalizes to reactive systems naturally.
More data analysis mechanisms are admissible for sophisticated formal
privacy analysis through our techniques.
}

@@

\noindent
\paragraph{Contributions.} Our main contributions are threefold. 
\begin{enumerate}
\item We introduce the logic $\dpCTLstar$ for reasoning about properties in
differential privacy. The logic is able to to express classical
differentially private properties;
%Complex properties such as those involving infinite behaviors can
%also be expressed; 
\item We model several differentially private mechanisms in Markov chains or Markov decision processes; and
\item We show that the model checking
problem for Markov chains is standard. For Markov decision processes, we show that it is undecidable.
\end{enumerate}
